Global study of well-being and internet adoption | Online analysis supplement

### 2021-11-22 03-model-estimate.Rmd
Split modelling to two files

# Time trends

@@

# MCMC settings
options(brms.backend = "cmdstanr")
hmc <- list(
  chains = 4,
  cores = 4,
  threads = 3,
  iter = 1500, 
  warmup = 750,
  refresh = 10,
  adapt_delta = .90, 
  max_treedepth = 10
)

# Parallelisation options: I have hard coded 12 cores here
# n_cores <- 12
# n_workers <- floor(n_cores / 4) # 4 chains always
# cluster <- new_cluster(n_workers)
# cluster_library(cluster, c("stringr", "dplyr"))

# Run this to delete fitted models and refit
# unlink(list.files(pattern = "output/brm-.*\\.rds"))
```

@@

### Figure: Timecourses and parameters

```{r}
tmp_coefs <- fits_1c %>% 
  mutate(
    out = map(
      fit, 
      ~hypothesis(
        .x, 
        # Rescale coefficient here to years on %
        # Note self harm doesnt work now
        c(str_glue("val_year = 0")),
        scope = "coef",
        group = "region"
      )$hypothesis
    )
  ) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  mutate(across(where(is.numeric), ~round(., 2))) %>% 
  mutate(result = str_glue('**{Estimate}**<br>[{CI.Lower}, {CI.Upper}]')) %>% 
  select(cause, region = Group, result)
tmp_coefs
pseudo_gwp <-
  bind_rows(
    distinct(gbd, region, year, val = NA) %>% 
      crossing(
        cause = fct_inorder(
          c(
            "Negative\nexperiences", 
            "Positive\nexperiences", 
            "Life\nsatisfaction"
          )
        )
      )
  )

# Stack data sets
# Assume all outcomes are proportions (0-1)
dat_all <- bind_rows(
  pseudo_gwp,
  gbd,
  itu %>% mutate(cause = "Internet", val = internet)
) %>% 
  mutate(cause = fct_inorder(cause)) %>% 
  group_by(region, country, year, cause) %>% 
  summarise(mean = mean(val, na.rm = TRUE)) %>% 
  ungroup()

# Draw small multiples figure, outcomes x regions
dat_all %>% 
  mutate(year = make_date(year)) %>% 
  ggplot(aes(year, mean)) +
  scale_y_continuous(
    "Value",
    labels = function(x) percent(x, .01) %>% str_remove("\\.00"),
    breaks = pretty_breaks()
  ) +
  scale_x_date(
    "Year",
    breaks = make_date(c(2000, 2005, 2010, 2015)), 
    date_labels = "'%y", 
    expand = expansion(.02)
  ) +
  geom_line(
    aes(group = interaction(country, cause)), 
    alpha = .25, size = .4, color = "dodgerblue1"
  ) +
  stat_summary(
    fun = mean, geom = "line", 
    size = .6, color = "dodgerblue4"
  ) +
  facet_grid(cause~region, scales = "free_y") +
  theme(
    legend.position = "none",
    panel.spacing = unit(3, "pt")
  ) +
  geom_richtext(
    data = tmp_coefs %>% mutate(cause = fct_inorder(str_to_title(cause))),
    aes(x = make_date(2000), y = Inf, label = result),
    size = 2.4, vjust = .9, hjust = 0.05, color = "black",
    family = Font,
    fill = NA, label.color = NA
  )
```

### Figure: Correlations

Calculate country coefficients and show scatterplots

```{r time-correlation-anxiety}
# Function to wrangle draws at multiple levels and create scatterplots + ellipses
draw_scatterplot <- function(fit, beta, r_region, r_country, sd_country, cor_country) {
  # Get average and region-specific slopes
  r_region_draws <- fit %>% 
    spread_draws(
      b_val_year, 
      b_internet_year,
      r_region__val[region, term],
      r_region__internet[region, term]
    ) %>% 
    filter(term == "year") %>% 
    ungroup() %>% 
    # Fix names by replacing dot with line break
    mutate(region = str_replace(region, "\\.", "\n"))

  # Then country
  r_country_draws <- fit %>% 
    spread_draws(
      r_country__val[country, term],
      r_country__internet[country, term]
    ) %>% 
    filter(term == "year") %>% 
    ungroup() %>% 
    # Fix names by taking labels from original names
    mutate(
      country = factor(
        country, 
        labels = sort(unique(fit$data$country))
      )
    ) %>% 
    left_join(distinct(dat, region, country))

  # Calculate country coefficients
  country_coefs <- left_join(r_country_draws, r_region_draws) %>% 
    mutate(
      internet = b_internet_year + r_region__internet + r_country__internet,
      outcome = b_val_year + r_region__val + r_country__val
    ) %>% 
    group_by(country, region) %>% 
    median_qi(internet, outcome)

  # Create country ellipse data
  country_ellipse <- spread_draws(
    fit,
    b_val_year, 
    b_internet_year,
    sd_country__val_year,
    sd_country__internet_year,
    cor_country__internet_year__val_year, 
    ndraws = 100
  ) %>% 
    rename(
      cor = cor_country__internet_year__val_year, 
      sd_outcome = sd_country__val_year,
      sd_i = sd_country__internet_year,
      b_outcome = b_val_year,
      b_i = b_internet_year
    ) %>% 
    mutate(
      e = pmap(
        list(cor, sd_outcome, sd_i, b_outcome, b_i), 
        function(cor, sd_a, sd_i, b_a, b_i) 
          ellipse(
            cor, 
            scale = c(sd_i, sd_outcome), 
            centre = c(b_i, b_outcome), 
            level = .90
          ) %>% 
          as_tibble()
      )
    ) %>% 
    unnest(e)

  # Correlation
  country_cor_result <- spread_draws(fit, cor_country__internet_year__val_year) %>% 
    median_qi(.width = .90) %>% 
    mutate(across(where(is.numeric), ~round(., 2))) %>%
    {str_glue("{.[,1]} [{.[,2]}, {.[,3]}]")}

  # Plot
  country_coefs %>% 
    ggplot(aes(internet, outcome)) +
    geom_path(
      data = country_ellipse,
      aes(x = x, y = y, group = .draw), 
      alpha = .2, size = .1, col = "dodgerblue1"
    ) +
    geom_vline(xintercept = 0, lty = 2, size = .2) +
    geom_hline(yintercept = 0, lty = 2, size = .2) +
    geom_point(col = "dodgerblue1", size = 1, alpha = .5) +
    scale_x_continuous(
      "Internet trend",
      breaks = pretty_breaks()
    ) +
    scale_y_continuous(
      breaks = pretty_breaks()
    ) +
    annotate(
      "text",
      x = Inf, y = Inf, label = country_cor_result,
      size = 4, vjust = 1.5, hjust = 1.05, color = "dodgerblue4",
      family = Font
    )
}
```

```{r}
p1 <- draw_scatterplot(fits_1c$fit[[1]]) + labs(y = "Anxiety trend")
p2 <- draw_scatterplot(fits_1c$fit[[2]]) + labs(y = "Depression trend")
p3 <- draw_scatterplot(fits_1c$fit[[3]]) + labs(y = "Self-harm trend")
```

```{r}
(wrap_elements(grid::textGrob('Negative experiences\nTBD')) |
   wrap_elements(grid::textGrob('Positive experiences\nTBD')) |
   wrap_elements(grid::textGrob('Life satisfaction\nTBD'))) /
  (p1 | p2 | p3)
```

@@

### Forest plots

figure of coefficients forest plot

```{r}
tmp <- fits_2c %>% 
  mutate(
    out = map(
      fit, 
      ~spread_draws(.x, b_i1_cw, r_region[region, term] | region) %>% 
        filter(term == "i1_cw") %>% 
        ungroup() %>% 
        select(-term) %>% 
        rename(Average = b_i1_cw) 
    )
  ) %>% 
  select(-fit) %>% 
  unnest(out)

names(tmp) <- str_replace(names(tmp), "\\.", " ")

tmp_text <- tmp %>% 
  group_by(cause) %>% 
  summarise(parameters(cur_data()[-c(1:3)], test = "pd")) %>% 
  rename(region = Parameter) %>% 
  mutate(is_average = region == "Average") %>% 
  mutate(region = if_else(region=="Average", "**Average**", region)) %>% 
  mutate(across(Median:CI_high, ~format(round(., 2), nsmall = 2)))

p_tmp <- tmp %>% 
  mutate(across(Africa:`Southern Asia`, ~.+Average)) %>% 
  pivot_longer(Average:`Southern Asia`, names_to = "region") %>% 
  mutate(region = if_else(region=="Average", "**Average**", region)) %>% 
  mutate(region = fct_relevel(factor(region), "**Average**")) %>% 
  mutate(cause = factor(cause)) %>% 
  mutate(is_average = region == "**Average**") %>% 
  ggplot(aes(value, region, fill = is_average, alpha = is_average)) +
  scale_fill_manual(values = c("dodgerblue1", "dodgerblue4")) +
  scale_alpha_manual(values = c(.4, .8)) +
  geom_vline(xintercept = 0, lty = 2, size = .2) +
  scale_shape_manual(values = c(21, 19)) +
  scale_y_discrete(expand = expansion(c(0.01, .1))) +
  scale_x_continuous(
    "Magnitude of association between internet adoption and well-being", 
    breaks = pretty_breaks(), 
    expand = expansion(c(.1, .5))
  ) +
  stat_halfeye(
    point_interval = NULL, normalize = "panels", 
    height = .75, adjust = 1.5
  ) +
  stat_pointinterval(
    alpha = 1, .width = c(.9),
    interval_size = 1,
    point_size = 1.5,
    fill = "white",
    aes(shape = after_stat(xmin > 0 | xmax < 0)),
    position = position_nudge(0, -.025)
  ) +
  geom_text(
    data = tmp_text, alpha = 1, size = 2.75, hjust = 1, vjust = -0.2,
    aes(
      x = Inf,
      label = str_glue("{Median} [{CI_low}, {CI_high}]")
    )
  ) +
  facet_wrap("cause", nrow = 1, scales = "free_x") +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_markdown(),
    legend.position = "none"
  )

(wrap_elements(grid::textGrob('Negative experiences\nTBD')) |
    wrap_elements(grid::textGrob('Positive experiences\nTBD')) |
    wrap_elements(grid::textGrob('Life satisfaction\nTBD'))) /
  p_tmp
```

### Figure: Between country effects

```{r}
tmp <- dat %>% 
  group_by(region, country, cause) %>% 
  summarise(
    internet = mean(internet, na.rm = TRUE),
    val = mean(val, na.rm = TRUE)
  ) 
tmp %>% 
  group_by(cause, region) %>% 
  summarise(tidy(lm(scale(val) ~ scale(internet), data = cur_data()))) %>% 
  filter(term != "(Intercept)") 
tmp %>% 
  ggplot(aes(internet, val)) +
  geom_point() +
  stat_smooth(method = "lm", se = F) +
  facet_grid(cause~region, scales = "free_y")

plot(conditional_effects(fits_2c$fit[[3]], "i1_cb"), points = TRUE)
x <- conditional_effects(fits_2c$fit[[3]], "i1_cb")[[1]]

x <- fits_2c$fit[[3]]
y <- fitted(x)
as_tibble(cbind(x$data, y)) %>% 
  arrange(desc(Estimate))
```

### Sex and Age

```{r}
tmp <- fits_2c %>% 
  mutate(
    out = map(
      fit, 
      ~emtrends(.x, var = "i1_cw", by = c("sex"), at = list(age1=0, age2=0))
    )
  )
tmp$out[[1]] %>% summary
```

Calculate coefficients from contrast coded factors

```{r}
x <- fits_2c$fit[[1]]
y <- emtrends(x, var = "i1_cw", by = c("age", "sex"))
gather_emmeans_draws(y) %>% mean_qi()

contrasts(x$data$sex)
contrasts(x$data$age)
h <- c(
  `Female_10 to 14` = 
    "i1_cw + sex1:i1_cw*1 + age1:i1_cw*1 + age2:i1_cw*0 = 0",
  `Female_15 to 19` = 
    "i1_cw + sex1:i1_cw*1 + age1:i1_cw*0 + age2:i1_cw*1 = 0",
  `Female_20 to 24` = 
    "i1_cw + sex1:i1_cw*1 + age1:i1_cw*-1 + age2:i1_cw*-1 = 0",
  `Male_10 to 14` = 
    "i1_cw + sex1:i1_cw*1 + age1:i1_cw*1 + age2:i1_cw*0 = 0",
  `Male_15 to 19` = 
    "i1_cw + sex1:i1_cw*1 + age1:i1_cw*0 + age2:i1_cw*1 = 0",
  `Male_20 to 24` = 
    "i1_cw + sex1:i1_cw*1 + age1:i1_cw*-1 + age2:i1_cw*-1 = 0"
)
tmp_avg <- hypothesis(x, h)
tmp_region <- hypothesis(x, h, scope = "coef", group = "region")
```

```{r}
# Get some stuff and transform
get_draws <- function(x) {
  draws_region <- spread_draws(
    x, 
    b_i1_cw, `b_sex1:i1_cw`, `b_age1:i1_cw`, `b_age2:i1_cw`,
    r_region[region, term] | term
  ) %>% 
    mutate(
      `10 to 14` = b_i1_cw + i1_cw + 
        `age1:i1_cw`*1 + `b_age1:i1_cw`*1 +
        `age2:i1_cw`*0 + `b_age2:i1_cw`*0,
      `15 to 19` = b_i1_cw + i1_cw + 
        `age1:i1_cw`*0 + `b_age1:i1_cw`*0 +
        `age2:i1_cw`*1 + `b_age2:i1_cw`*1,
      `20 to 24` = b_i1_cw + i1_cw + 
        `age1:i1_cw`*-1 + `b_age1:i1_cw`*-1 +
        `age2:i1_cw`*-1 + `b_age2:i1_cw`*-1,
      Female = b_i1_cw + i1_cw + `sex1:i1_cw`*1 + `b_sex1:i1_cw`*1,
      Male = b_i1_cw + i1_cw + `sex1:i1_cw`*-1 + `b_sex1:i1_cw`*-1,
      Age = `20 to 24` - `10 to 14`,
      Sex = Female - Male
    ) %>% 
    select(region, `10 to 14`:Sex)

  draws_avg <- spread_draws(
    x, 
    b_i1_cw, `b_sex1:i1_cw`, `b_age1:i1_cw`, `b_age2:i1_cw`,
  ) %>% 
    mutate(
      `10 to 14` = b_i1_cw + `b_age1:i1_cw`*1 + `b_age2:i1_cw`*0,
      `15 to 19` = b_i1_cw + `b_age1:i1_cw`*0 + `b_age2:i1_cw`*1,
      `20 to 24` = b_i1_cw + `b_age1:i1_cw`*-1 + `b_age2:i1_cw`*-1,
      Female = b_i1_cw + `b_sex1:i1_cw`*1,
      Male = b_i1_cw + `b_sex1:i1_cw`*-1,
      Age = `20 to 24` - `10 to 14`,
      Sex = Female - Male
    ) %>% 
    select(`10 to 14`:Sex) %>% 
    mutate(region = "**Average**")

  draws <- bind_rows(
    "regions" = draws_region,
    "average" = draws_avg, 
    .id = "Level"
  ) %>% 
    mutate(region = str_replace(region, "\\.", "\n")) %>% 
    ungroup()
  draws
}

draws <- fits_2c %>% 
  mutate(out = map(fit, get_draws)) %>% 
  select(-fit) %>% 
  unnest(out)

# Summaries of differences
draws_sum <- draws %>% 
  select(cause, region, Age, Sex) %>% 
  group_by(cause, region) %>% 
  summarise(
    describe_posterior(
      cur_data(), centrality = "mean", test = "pd"
    )
  ) %>% 
  rename(panel = Parameter) %>% 
  mutate(across(where(is.numeric), ~round(., 2)))

draws %>% 
  select(cause, Level, region, c(`10 to 14`, `20 to 24`, Female, Male)) %>% 
  pivot_longer(-c(region, Level, cause), names_to = "Group") %>% 
  mutate(panel = if_else(str_detect(Group, "ale"), "Sex", "Age")) %>% 
  mutate(
    side = if_else(Group %in% c("10 to 14", "Female"), "top", "bottom")
  ) %>% 
  mutate(region = str_replace(region, "\\.", " ")) %>% 
  mutate(region = fct_relevel(region, "**Average**")) %>% 
  ggplot(aes(value, region, col = Group, fill = Group)) +
  scale_alpha_manual(values = c(.6, .3)) +
  geom_vline(xintercept = 0, lty = 2, size = .2) +
  scale_shape_manual(
    values = c(21, 19), guide = "none"
  ) +
  scale_y_discrete(expand = expansion(c(0.1, 0.1))) +
  scale_x_continuous(
    "Magnitude of association between internet adoption and well-being", 
    breaks = pretty_breaks(), 
    expand = expansion(c(0, .05))
  ) +
  stat_halfeye(
    point_interval = NULL, normalize = "panels", 
    height = .75, adjust = 1.5,
    aes(side = side, group = side, alpha = Level), 
    show.legend = FALSE
  ) +
  stat_pointinterval(
    alpha = 1, .width = c(.9),
    interval_size = .33,
    point_size = 1.25,
    fill = "white",
    aes(shape = after_stat(xmin > 0 | xmax < 0), group = side),
    position = position_dodgev(.25), key_glyph = draw_key_point
  ) +
  geom_richtext(
    data = draws_sum %>% 
      mutate(Group = NA) %>% 
      filter(cause %in% c("Anxiety", "Depression", "Selfharm")),
    aes(
      x = Inf, 
      label = str_glue("*d* = {Mean} (p<sup>d</sup> = {percent(pd, 1)})")
    ),
    size = 2.5, hjust = 1,
    family = Font,
    fill = NA, label.color = NA,
    show.legend = FALSE
  ) +
  facet_grid2(panel ~ cause, scales = "free_x", axes = "x") +
  guides(
    alpha = "none"
  ) +
  theme(
    legend.position = "right",
    axis.title.y = element_blank(),
    axis.text.y = element_markdown()
  )
```

### Fitted lines

Plot of fitted lines

```{r}
# Prediction dataframe: Min and max vals are enough for line
x_country <- fits_2c$fit[[1]]$data %>% 
  group_by(country) %>% 
  mutate(
    xmin = min(i1_cw, na.rm = TRUE), 
    xmax = max(i1_cw, na.rm = TRUE)
  ) %>% 
  filter(i1_cw %in% c(xmin, xmax)) %>% 
  mutate(year = 0) %>% 
  ungroup()

pred_country <- fits_2c %>% 
  mutate(
    pred_country = map(
      fit, ~epred_draws(
        .x, 
        x_country,
        ndraws = 100
      ) %>% 
        group_by(region, country, i1_cw) %>% 
        mean_qi(.epred)
    )
  )

x_region <- x_country %>% 
  distinct(region, sex, age, i1_cb) %>% 
  group_by(region) %>% 
  mutate(i1_cb = mean(i1_cb, na.rm = T)) %>% 
  crossing(i1_cw = seq(-.4, .4, length = 10)) %>% 
  mutate(year = 0, se = 1)

pred_region <- fits_2c %>% 
  mutate(
    pred_region = map(
      fit, ~epred_draws(
        .x, 
        x_region,
        re_formula = 
          ~(year * sex + year * age + i1_cw * sex + i1_cw * age | region),
        ndraws = 100
      ) %>% 
        group_by(region, i1_cw) %>% 
        mean_qi(.epred)
    )
  )

pred_region <- pred_region %>% 
  select(-fit) %>%
  unnest(pred_region) %>%
  mutate(cause = fct_rev(fct_inorder(factor(cause))))

pred_country %>% 
  select(-fit) %>% 
  unnest(pred_country) %>% 
  bind_rows(select(pseudo_gwp, region, cause)) %>% 
  mutate(cause = fct_rev(fct_inorder(factor(cause)))) %>% 
  ggplot(aes(i1_cw, .epred)) +
  scale_x_continuous(
    "Previous year's internet adoption relative to country average in 2000-2020",
    breaks = c(-.35, 0, .35),
    labels = function(x) percent(x, 1)
  ) +
  scale_y_continuous(
    "Estimated rate or scale score",
    breaks = pretty_breaks()
  ) +
  geom_blank() +
  geom_ribbon(
    data = pred_region,
    aes(ymin = .lower, ymax = .upper),
    alpha = .15, fill = "dodgerblue4"
  ) +
  geom_line(
    data = pred_region,
    col = "dodgerblue4", size = .6
  ) +
  geom_line(
    aes(group = country), 
    size = .4, alpha = .35, col = "dodgerblue1"
  ) +
  # geom_text(
  #   data = tmp_text %>% 
  #     filter(region != "**Average**") %>% 
  #     bind_rows(select(pseudo_gwp, region, cause)) %>% 
  #     mutate(cause = fct_rev(fct_inorder(factor(cause)))) %>% 
  #     mutate(region = str_replace(region, " ", "\n")) %>% 
  #     mutate(result = paste(Median)),
  #   aes(x = -Inf, y = Inf, label = result),
  #   size = 2.4, vjust = .9, hjust = 0.05, color = "black",
  #   family = Font,
  #   fill = NA, label.color = NA
  # ) +
  facet_grid(cause~region, scales = "free_y") +
  theme(panel.spacing = unit(2, "pt"))
```

>>>>

# Model fitting

@@

# MCMC settings
options(brms.backend = "cmdstanr")
hmc <- list(
  chains = 4,
  cores = 4,
  threads = 3,
  iter = 1500, 
  warmup = 750,
  refresh = 10,
  adapt_delta = .90, 
  max_treedepth = 10
)
```

@@

@@
